---
title: "Getting Started with problmsolver"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with problmsolver}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

`problmsolver` is an R wrapper over Python `problm_solver` via `reticulate`.
It provides a small, explicit R API for model loading, querying, adjusted
sampling, and dataset utilities.

## Backend setup

### Managed backend (recommended)

```{r eval = FALSE}
library(problmsolver)

# Creates/updates a managed virtualenv and installs `problm-solver`
ps_backend_setup()
ps_use_backend_env()
```

### From a local clone (no PyPI required)

```{r eval = FALSE}
# If you already cloned probLM-solver locally:
ps_backend_setup_local("/abs/path/to/probLM-solver")
```

### Auto-create on first configure

```{r eval = FALSE}
ps_configure(envname = "r-problmsolver", auto_create = TRUE)
```

## Minimal model example

```{r eval = FALSE}
model <- ps_model(
  fname = "/path/to/model.gguf",
  context = "Why is the sky blue?",
  n_ctx = 4096L,
  logits_all = TRUE,
  n_gpu_layers = 999L
)

ps_query(model)
```

## Sampling examples

```{r eval = FALSE}
# Low-temperature
low_temp <- ps_sample_low_temp(alpha = 2.0)
out_low <- ps_generate_adjusted(
  model = model,
  top_k = 8L,
  top_p = 0.9,
  adjust_fn = low_temp,
  max_tokens = 128L
)

# Power distribution + Metropolis branch sampling
metro <- ps_metropolis_sampler(equil_branches = 1L, max_branches = 10L)
power <- ps_sample_power_dist(alpha = 2.0, lookahead_depth = 10L, branch_sampler = metro)
out_power <- ps_generate_adjusted(
  model = model,
  top_k = 8L,
  top_p = 0.9,
  adjust_fn = power,
  max_tokens = 128L
)
```

## Custom sampling function example

```{r eval = FALSE}
my_adjust_r <- function(ctx) {
  out <- ctx$token_probs
  if (' the' %in% names(out)) {
    out[' the'] <- out[' the'] + 0.25
  }
  out
}

out_custom <- ps_generate_adjusted(
  model = model,
  top_k = 8L,
  top_p = 0.9,
  adjust_fn = ps_r_adjust_fn(my_adjust_r),
  max_tokens = 128L
)
```

## Dataset utility example

```{r eval = FALSE}
math500 <- ps_get_math500()
problems <- ps_get_problems_math500()
head(problems)
```

## Notes for robust usage

- Set `logits_all = TRUE` when using adjusted generation.
- Use `ps_reset_module()` if you switch Python environments mid-session.
- Keep one backend env per project for reproducibility.
